---
title: "Text Mining Sentiment from Hotel Reviews"
author: "Ian Jeffries"
date: "5/17/2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The following R Markdown outlines the steps used by Ian Jeffries to mine sentiment from over 21,000 hotel reviews on resorts located in the Republic of Maldives, a South Asian country located in the Indian Ocean. 

## Install necessary packages

The following code will install the packages used in the project:

```{r packages, message=FALSE, warning=FALSE}
#list of packages used
packages <- c("tm", "wordcloud", "lubridate", "SnowballC")

#check to see if package is already installed, if not, install
for(p in packages){
  if(!require(p, character.only = TRUE)) {
    install.packages(p)
    library(p, character.only = TRUE)
  } 
}
```
## Import the Hotel Review Dataset

Import the data from my github page

```{r data import}
path <- "https://raw.githubusercontent.com/ianjeffries/review-sentiment-analysis/master/data/maldives_hotel_reviews.csv"
reviews <- read.csv(path, header = TRUE, stringsAsFactors = FALSE)
head(reviews)
str(reviews)
```

## Cleaning the Data

The data can now be prepared for analysis.

```{r data cleaning}
#change date to date format
reviews$Review.Date <- dmy(reviews$Review.Date)

#fix "invalid multibyte" error
for (i in c(1, 2, 5:8)) {
  reviews[ ,i] <- enc2utf8(reviews[,i])
}

#remove rows without review
reviews <- reviews[!reviews$Review.Heading=="", ]

#add in review number
reviews$review.number <- 1:21039
```

```{r}
```


```{r}
```
